High-Performance Language Primitives for Distributed Architectures using Asynchronous Events

### 2012-07-23 doc/asplos2013/impl.tex
space compaction, including the nuking of a duplicate paragraph

Barriers are implemented as an extension of events.  In place of an event's triggered/untriggered state,
a barrier's owner tracks the number of outstanding arrivals.  When the arrival count 
reaches zero, the barrier is considered to have triggered.  Remote nodes send {\em barrier update}
active messages containing an adjustment value that is positive for {\tt alter\_arrival\_count} operations 
and negative for {\tt arrive} operations.  A simplified version of vector clocks\cite{Fidge1998} is used
to detect the race condition in which the barrier owner receives an {\tt arrive} operation before the
corresponding {\tt alter\_arrival\_count}.  When detected, the decrement of the arrival count 
is delayed to avoid a spurious triggering of the barrier.

@@

When a lock request is made, an event is created to track when the grant occurs.  The current node then
examines its copy of the lock data structure to determine if it is the owner.
%(creating it if this the first operation for that lock on that node).  
If the current node is the owner and the lock isn't
held, the lock is granted immediately and the event is triggered.  If the lock is held, the event is added
to the list of local waiters.  Note that the event associated with
the lock request is the only data that must be stored.
%(This is the only thing that is stored.  The actual ``requestor'' of the lock
%is implicitly captured in which operations are dependent on the lock grant event.)  
If the current node isn't the owner, a {\em lock request} active message is
sent to the most recently known owner.  If the receiving node is no
longer the owner, it will forward the message on to the next node known
to be the owner of the lock.
%(If that information proves to be stale, the request is forwarded
%by the receiving node, eventually catching up to the current owner.)  
If the current owner's status shows
the lock is currently held, the bit for the requesting node is set in remote waiters bitmask.  If the lock is
not held, the ownership of the lock is given to the requesting node via a {\em lock transfer} active
message.  A lock transfer message includes the bitmask of remote waiters and an up-to-date copy 
of the lock's payload.  (The inclusion of the payload in the active message is the reason for the 
4KB limit on a payload's size.)

@@

The unfairness that results from a lock favoring local waiters over remote waiters is intentional.  When the
contention on a lock is high (the only time when the question of fairness is relevant), the latencies
involved in transferring a lock between nodes can be the limiter on throughput.  By minimizing the number
of lock transfers, the throughput on the lock is maximized.  This effect will be demonstrated in
Section~\ref{subsec:lockmicro}.

>>>>

Barriers are implemented as an extension of events.  In place of an event's triggered/untriggered state,
a barrier's owner tracks the number of outstanding arrivals.  When the arrival count 
reaches zero, the barrier is considered to have triggered.  Remote nodes send {\em barrier update}
active messages containing an adjustment value that is positive for {\tt alter\_arrival\_count} operations 
and negative for {\tt arrive} operations.  A simplified version of vector clocks\cite{Fidge1998} is used
to detect the race condition in which the barrier owner receives an {\tt arrive} operation before the
corresponding {\tt alter\_arrival\_count}.  When detected, the decrement of the count 
is delayed to avoid a spurious triggering of the barrier.

@@

When a lock request is made, an event is created to track when the grant occurs.  The current node then
examines its copy of the lock data structure to determine if it is the owner.
%(creating it if this the first operation for that lock on that node).  
If the current node is the owner and the lock isn't
held, the lock is granted immediately and the event is triggered.  If the lock is held, the event is added
to the list of local waiters.  Note that the event associated with
the lock request is the only data that must be stored.
%(This is the only thing that is stored.  The actual ``requestor'' of the lock
%is implicitly captured in which operations are dependent on the lock grant event.)  
If the current node isn't the owner, a {\em lock request} active message is
sent to the most recently known owner.  If the receiving node is no
longer the owner, it forwards the message on to the node it has recorded as the owner.
%(If that information proves to be stale, the request is forwarded
%by the receiving node, eventually catching up to the current owner.)  
If the current owner's status shows
the lock is currently held, the bit for the requesting node is set in remote waiters bitmask.  If the lock is
not held, the ownership of the lock is given to the requesting node via a {\em lock transfer} active
message.  A lock transfer message includes the bitmask of remote waiters and an up-to-date copy 
of the lock's payload.  (The inclusion of the payload in the active message is the reason for the 
4KB size limit.)

@@

The unfairness that results from a lock favoring local waiters over remote waiters is intentional.  When the
contention on a lock is high (the only time the question of fairness is relevant), the latencies
involved in transferring a lock between nodes can be the limiter on throughput.  By minimizing the number
of lock transfers, the throughput on the lock is maximized.  This effect will be demonstrated in
Section~\ref{subsec:lockmicro}.

### 2012-07-23 doc/asplos2013/interface.tex
space compaction, including the nuking of a duplicate paragraph

\lstset{
  captionpos=b,
  language=C++,
  basicstyle=\scriptsize,
  numbers=left,
  numberstyle=\tiny,
  columns=fullflexible,
  stepnumber=1,
  escapechar=\#,
  keepspaces=true,
  %literate={<}{{$\langle$}}1 {>}{{$\rangle$}}1,
  %morekeywords={region,coloring,partition,spawn,disjoint,aliased},
  %deletekeywords=float,
}

@@

Although {\tt UserEvent}s and {\tt Barrier}s are created and triggered differently,
any operation that expresses a dependence on an {\tt Event} can transparently use
a {\tt UserEvent} or {\tt Barrier} as well.

@@

\begin{lstlisting}[float={t},label={lst:procapi},caption={Processor Interface and Example.}]
class Processor {
public:
  const unsigned id;
  typedef unsigned TaskFuncID;
  typedef void (*TaskFuncPtr)
            (const void *args,size_t arglen,Processor p);
  typedef map<TaskFuncID, TaskFuncPtr> TaskIDTable;

@@

  \node (t4) at (6.4,0) [shape=rectangle,draw] {\begin{tabular}{c}{\bf run}\\TASK\_D\end{tabular}};
  \draw [->] (t4) to node[auto] {ed} (7.9,0);
  \draw [->] (5.4,0) to (t4.west);
  %
\end{tikzpicture}
}
\caption{Processor Example Event Graph\label{fig:procevents}}
\end{figure}

@@

  \draw [->,dotted,thick] (-1,1) to (t1.135);
  \draw [->,dotted,thick] (t3.45) to (7.4,1);
\end{tikzpicture}
}
\caption{Lock Example Event Graph\label{fig:lockevents}}
\end{figure}

@@

\end{tikzpicture}
}
\caption{Reduction Example Event Graph\label{fig:reducevents}}
\end{figure}

>>>>

\lstset{
  captionpos=b,
  language=C++,
  basicstyle=\scriptsize,
  numbers=left,
  numberstyle=\tiny,
  columns=fullflexible,
  stepnumber=1,
  escapechar=\#,
  keepspaces=true,
  belowskip=-10pt,
  %literate={<}{{$\langle$}}1 {>}{{$\rangle$}}1,
  %morekeywords={region,coloring,partition,spawn,disjoint,aliased},
  %deletekeywords=float,
}

@@

Although UserEvents and Barriers are created and triggered differently,
operations that express dependences on Events can transparently use
UserEvents or Barriers as well.

@@

\begin{lstlisting}[float={t},label={lst:procapi},caption={Processor Interface and Example.},belowskip=0pt]
class Processor {
public:
  const unsigned id;
  typedef unsigned TaskFuncID;
  typedef void (*TaskFuncPtr)
            (const void *args,size_t arglen,Processor p);
  typedef map<TaskFuncID, TaskFuncPtr> TaskIDTable;

@@

  \node (t4) at (6.4,0) [shape=rectangle,draw] {\begin{tabular}{c}{\bf run}\\TASK\_D\end{tabular}};
  \draw [->] (t4) to node[auto] {ed} (7.9,0);
  \draw [->] (5.4,0) to (t4.west);
  %
\end{tikzpicture}
}
\vspace{-2mm}
\caption{Processor Example Event Graph\label{fig:procevents}}
\vspace{-4mm}
\end{figure}

@@

  \draw [->,dotted,thick] (-1,1) to (t1.135);
  \draw [->,dotted,thick] (t3.45) to (7.4,1);
\end{tikzpicture}
}
\vspace{-2mm}
\caption{Lock Example Event Graph\label{fig:lockevents}}
\vspace{-4mm}
\end{figure}

@@

\end{tikzpicture}
}
\vspace{-6mm}
\caption{Reduction Example Event Graph\label{fig:reducevents}}
\vspace{-4mm}
\end{figure}
